PMV model comparison

### 2025-01-28 Manuscript/src/response-letter.tex

            \response{
                Thank you for this very valuable comment.
                The ASHRAE Standard defines comfort as PMV between -0.5 and 0.5. 
                The European standard proposes four classes. 
                We agree that limiting the applicability of the PMV model to the range from \numrange{-.5}{.5} may, in some instances, increase the building operation energy consumption.
                However, in some conditions that is not necessarily the case, for example, by providing individual control over the air speed.
                For example, in Figure 2 of the manuscript, we show that the PMV predicts that occupants (met=1.2 and \ac{clor}=0.5) are comfortable at \ac{tdb}=27.7C when \ac{vr}~=~0.4 m/s.
                A \ac{tdb}=27.7C is higher than the average indoor air temperature in offices in the US, Europe, Australia and Singapore which is generally in the range of \qtyrange{22}{24}{\celsius}.
                We have now included the following paragraph in the conclusion of the manuscript to address this issue: \textit{``We understand that this may raise the concern that limiting the \ac{pmv} to $\lvert \textrm{PMV}\lvert \leq 0.5$ would increase the building energy consumption.
                However, this is not always the case.
                As shown in Figure 2 increasing \ac{vr} to a modest \qty{0.4}{\m\per\s}, which is achievable with any standard electric fan, is sufficient in an office setting to extend the upper boundary of comfort region to \ac{tdb}~=~\qty{27.7}{\celsius}.
                This is the upper \ac{tdb} value estimated by the \ac{pmv} to keep a person within the comfort region while wearing a typical office attire \ac{clor}~=~\qty{0.5}{clo} and performing office tasks with \ac{met}~=~\qty{1.2}{met} with \ac{rh}~=~\qty{50}{\percent}.
                Hence, we would like to point out that the \ac{pmv} model should be used in a range where it is accurate, and then the designer should look for solutions that provide a more sustainable built environment.
                A field study in Singapore showed that allowing participants to control \ac{v} while simultaneously increasing \ac{tdb} from \qty{24}{\celsius} to \qty{26.5}{\celsius} reduced the energy consumption by \qty{32}{\percent}.''}
            }

            \item Some conclusion contents lack data and theoretical support.
            \response{
            We assume that the reviewer referred to the thermal preference conclusions. 
            We modified it by making our assumption, as described in the answer above, explicit and clear. 
            We modified the text as follows: "If we assume that selecting “warmer” or “cooler” on the thermal preference scale indicates thermal discomfort, then our results suggest that individuals who report being slightly warm' or slightly cool’ on the thermal sensation scale are thermally comfortable is incorrect. 
            This misinterpretation can lead to an underestimation of thermal discomfort. 
            In the \ac{db2} \qty{68}{\percent} of participants who reported to be `slightly warm' wanted to be `cooler' and \qty{54}{\percent} of participants who reported to be `slightly cool' wanted to be `warmer'.}

@@

        \textbf{Reviewer 3}
        The manuscript presents an excellent analysis of the discrepancies between the two versions of the PMV model and evaluates their prediction accuracy.
        The analysis is informative and relevant for researchers and professionals within thermal comfort.
        The structure of the manuscript and the presentation of the topic are well laid out.

        \response{
            Thank you for your positive feedback.
        }

        \begin{enumerate}
            \item I have only a specific few comments, which are listed below.
            However, my main concern deals with the way the records representing individuals in the comfort DB are used to first calculate PMV for individuals and then compare them with individual TSVs.
            As the authors write: ``The intended aim of the PMV model is not to accurately predict individual thermal response from participants.
            The PMV model was developed to predict the average thermal sensation of an undefined large group of occupants sharing the same environment.``
            But isn't this analysis relying on the calculation of individual TSV without considering that this is not what the model should be used for?
            The approach is the same across a wealth of studies that examine the accuracy of the model, but that does not make it more correct.
            I doubt that the DB can be reliably used to estimate the prediction accuracy, as there is no information on the mean TSV of groups of people exposed to the same thermal environment.

            \response{
                Thank you for your valuable comment.
                We agree with the reviewer that originally the \ac{pmv} was not intended to predict the thermal sensation of each participant and we have mentioned this in several sections of the paper, for example in Section 2.2.
                However, the PMV model is widely used in the scientific literature \textit{``in both singly occupied spaces and areas accommodating several hundred people''}.
                In Section 2.2 of the manuscript we have also added the following sentence \textit{``Partially because neither of the two standards specifies the minimum number of people needed to apply the model.''}.\\ 
                We are, however, aware of this issue and we have not neglected it.
                In the manuscript, we stated that \textit{``we subtracted the \ac{tsv} value from the \ac{pmv} and \ac{pmv-ce} values.
                These differences, also known as bias, quantify the success of the model in predicting \ac{tsv}.
                However, on their own, are a low-precision estimate of the overall accuracy of the model''.}
                We have also added the following sentence: \textit{``since the model is not expected to predict the exact \ac{tsv} of each participant}''.
                We believe that \textit{``If the \ac{pmv} or \ac{pmv-ce} formulations are bias-free, the distribution of any batch derived from these differences would have a mean value that is zero.
                The standard deviation would reflect the combined effect of the people's differences, any errors in the model formulation or, in the data collection method (accuracy or precision of the instrumentation used)''}.\\
                Finally, we would like to raise a concern we have about the vague definition of the PMV model in both the ISO 7730 and ASHRAE 55 standards.
                Stating that a model is intended to predict the average thermal sensation of an undefined large group of occupants is not a clear definition and, in principle, it makes it impossible to determine the accuracy of the model.
                At the same time, it does not help the user to understand when the model can be used and when it cannot.
                For example, can it be used to design a single office?
                What about when the space is occupied by two or three people?
                We believe that the Standards should provide a clearer definition of the range of applicability of the PMV model. 
                We have, therefore, added a sentence about this in the conclusion of the manuscript which reads as follows: \textit{``Finally, we would like to highlight that the current definition of the \ac{pmv} model which states that the model aims ``\ldots to predict the average thermal sensation of a large group of occupants'' is ambiguous.
                Both the \gls{55} and \gls{7730} standards should clarify the minimum number of occupants required to use the \ac{pmv} model.
                For example, can it be used for a space with only one occupant?''}
            }

@@

            \item Although less critical, another general comment is that it is not clear what is meant by low prediction accuracy, as is written already in the abstract -- when is the prediction accuracy low?
            We aim to estimate what people perceive, which seriously suffers from individual differences and many other challenges.
            What is the prediction accuracy of related models?
            Or is the low prediction accuracy an outcome of using the prediction model wrongly (see first general comment)?

            \response{
                Estimating perceived thermal comfort is inherently challenging due to individual differences and contextual factors, both of which significantly influence prediction accuracy. 
                There is no universal definition of high, medium, or low accuracy, as accuracy is context-dependent and should be evaluated based on the potential consequences of an error. 
                For example, in critical applications such as detecting cancer or making autonomous vehicle braking decisions, extremely high accuracy is essential. 
                In less consequential scenarios, such as determining whether an environment feels warm or cool, more flexibility is acceptable. 
                Negligible accuracy in any context can be considered equivalent to random guessing. 
                In our case, with seven thermal sensation categories, random guessing would result in an accuracy of approximately 14 percent. 
                Achieving a correct prediction one out of three times (around 33 percent) can be considered low accuracy because it is equivalent to a very simple model based solely on temperature achieved a similar level of accuracy (see Section 6 and Figure 7 of Cheung, Toby, Stefano Schiavon, Thomas Parkinson, Peixian Li, and Gail Brager, “Analysis of the Accuracy on PMV – PPD Model Using the ASHRAE Global Thermal Comfort Database II,” Building and Environment, 153, no. 15 (April 15, 2019): 205–17. https://doi.org/10.1016/j.buildenv.2019.01.055).
                One way to evaluate accuracy is to compare it with what could be achieved by other models. 
                For example, personal comfort models, now applied in smart thermostats like Nest or in car HVAC systems such as Tesla’s, have demonstrated accuracy levels exceeding 95 percent. 
                However, these models are not yet widely applicable to design scenarios. 
                Nevertheless, they tell us where we should aim. 
                We also drew on previous studies, such as the work by Humphreys and Nicol (2002), which provide a benchmark for evaluating accuracy. 
                In line with their approach, we calculated the bias of the model to determine whether it systematically over- or under-predicts participants’ thermal sensations.
                Individual differences in human subjects can be assumed to be random and distributed around the mean of a `typical person'.
                We have then added the following sentence to the manuscript: \textit{``This is a similar assumption to the one used by the \ac{pmv} model which ignores individual differences and calculates the average thermal sensation of a `typical' average person.''}
                Consequently, we believe that calculating the bias of the \ac{pmv} models is a valid approach to determine the overall accuracy of the model and that the low prediction accuracy is not an outcome of using the model wrongly.
            }

@@

            \item Ln. 103 - what is meant by formulation error?
            Unsuitable equations or constants that are wrong?

            \response{
                We apologize for the confusion.
                We meant both the formulation of the PMV model and the constants used in the model.
                We have added a sentence in the manuscript to clarify this point.
            }

            \item Sources of error.
            Order of importance -- is it the heat transfer calculation that is the most important -- not the attempt to calculate thermal sensation based on physical variables

            \response{We have combined the discussion of heat transfer issues and the intrinsic complexity of predicting people’s responses into a single paragraph in the paper, we did not rank them. 
            The title of this paragraph has been updated to: \textit{``Heat Balance Equations and Thermal Sensation Predictions.} 
            In the text, we emphasize: \textit{``The issues are reported in order of importance of what we believe may affect the accuracy of the model.} 
            This statement clarifies that we do not have quantifiable metrics to rank these factors; instead, the order reflects our expert judgment regarding their potential impact on model accuracy.}

@@

            \item ``Additionally, the PMV model erroneously assumes that the human body is always losing or gaining heat from its surrounding environment.
            In reality, under most indoor conditions, the body activates control mechanisms to maintain a stable core temperature [24].``
            Indeed, the thermoregulatory system maintains a stable body temperature, but the body still exchanges heat with the environment, so suggest revising this sentence.

            \response{
                We apologize for the mistake.
                We have revised the sentence to clarify this point and now reads as follows: \textit{``Additionally, the PMV model erroneously assumes that the human body is not capable of maintaining a stable core temperature.
                A \ac{pmv} value higher than \num{.5} or lower than \num{-.5} indicates that the hypothetical cylinder representing the human body is either gaining or losing heat, and consequently, it is getting warmer or colder, respectively.
                }
            }

@@

            \item P 23 ln 416.
            Calculation of heat loss by sweating.
            This equation applies to people in thermal comfort, which should probably be added to the statement, so heat lost due to sweating does not appear to be related only to met rate.

            \response{Agree, we modified the text as follows: \textit{``This amount of heat loss is included in the equation even if a person is in thermal neutrality (feeling comfortable) or has a negative heat gain (feeling cooler than neutral).}}

@@

            \item ``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent a TSV vote.'' This is wrong -- the PMV was never meant to represent a TSV vote, but the average of TSV votes of a large group of people.

            \response{
                Please see our answer above about the weak definition of what is "a large group of people" and the fact that standards allow the PMV model to be used in places occupied by only one person. 
                Nevertheless, we changed the sentence as follows: \textit{``The PMV model uses the overall heat losses or gains to calculate a PMV value which should represent the average thermal sensation of a large group of occupants.}
            }
